---
title: "eda_week_01"
author: "Doug Joubert"
date: "October 15, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Week 01: Exploratory Graphs (Part 1)
Read in some EPA Data
```{r}
library(tidyverse)
pollution <- read_csv("data/avgpm25.csv")
str(pollution)
```

I need to coererce the fips variable to character
```{r}
pollution$fips <- as.character(pollution$fips)
str(pollution)
head(pollution)
```

## One-dimension Summaries

### Boxplots

I guess I could do a simply boxplot of the pm25, with a summary table
```{r}
summary(pollution$pm25)
boxplot(pollution$pm25)
abline(h = 12) # this a/b line can be set to the value we are looking for, 12
```

### Histograms
So, this tells us that some are over the limit and some are under the limit. So, let us try a histogram

```{r}
hist(pollution$pm25, col = "light green")
rug(pollution$pm25) # need to run this as chunk, meaning the plot needs to be showing for rug to work
```

Let us experiment with changing the width of the bars in the histogram
```{r}
hist(pollution$pm25, col = "light green", breaks = 50) # makes the bars smaller, do not want the number to be too big or too small
```

You can also experiment with overlays on the histogram
```{r}
hist(pollution$pm25, col = "light green")
abline(v = 12, col = "red", lwd = 4) # Vertical line at 12
abline(v = median(pollution$pm25), col = "blue", lwd = 4) # Blue line at the median
```

### Barplot
A Barplot is another example, that can be use for categorical data
```{r}
barplot(table(pollution$region), col = "light blue", main = "Number of Counties in Each Region")
```

To get this work you can either create a table with the variable, or build the table into the barplot function

## Two-dimensional Summaries

### Multiple Overlays

### Boxplots with more that one Dimensions

This time (because we are displaying two boxplots), we modify the code that we previously used for the boxplot

```{r}
boxplot(pm25 ~ region, data = pollution, col = "light green")
```

This allows us to see that when compared to each region, the east region has a pm25 that is a little larger. Although, all of the extreme values are in the west region. The west region also has more spread, since the box is larger.

### Histograms with more that one Dimensions

```{r}
par(mfrow = c(2, 1), mar = c(4, 4, 2, 1))
hist(subset(pollution, region == "east") $pm25, col = "light green")
hist(subset(pollution, region == "west") $pm25, col = "light green")
```



### Scatterplots

```{r}
with(pollution, plot(latitude, pm25))
abline(h = 12, lwd = 2, lty = 2)
```

Of course, the instructor (again) explains absolutely nothing about the code that he uses in his example. So, I created the plot using ggplot


```{r}
# Or with the tidyverse
ggplot(data = pollution, mapping = aes(x = latitude, y = pm25)) +
  geom_point(mapping = aes(col = region)) +
  geom_smooth(method = lm, se=FALSE)
```

Then, I can use faceting to view the data using multiple scatterplots

```{r}
ggplot(data = pollution, mapping = aes(x = latitude, y = pm25)) +
  geom_point() +
  geom_smooth(method = lm, se = FALSE) +
  facet_wrap(~region)
```

# Plotting Systems in R
Basic example using base R

```{r}
library(datasets)
data("cars")
with(cars, plot(x = speed, y = dist))
```

Basic example using lattice package
```{r}
library(lattice)
state <- data.frame(state.x77, region = state.region)
xyplot(Life.Exp ~ Income | region, data = state, layout = c(columns = 4, rows = 1))
```

Basic example using base ggpolot
```{r}
library(tidyverse)
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) +
  geom_point()
```

