---
title: "eda_week_01"
author: "Doug Joubert"
date: "October 15, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Week 01: Exploratory Graphs (Part 1)
Read in some EPA Data
```{r}
library(tidyverse)
pollution <- read.csv("data/avgpm25.csv")
str(pollution)
```

I need to coererce the fips variable to character
```{r}
pollution$fips <- as.character(pollution$fips)
str(pollution)
head(pollution)
```

I guess I could do a simply boxplot of the pm25, with a summary table
```{r}
summary(pollution$pm25)
boxplot(pollution$pm25)
abline(h = 12) # this a/b line can be set to the value we are looking for, 12
```

So, this tells us that some are over the limit and some are under the limit. So, let us try a histogram

```{r}
hist(pollution$pm25, col = "light green")
rug(pollution$pm25) # need to run this as chunk, meaning the plot needs to be showing for rug to work
```

Let us experiment with changing the width of the bars in the histogram
```{r}
hist(pollution$pm25, col = "light green", breaks = 50) # makes the bars smaller, do not want the number to be too big or too small
```

You can also experiment with overlays on the histogram
```{r}
hist(pollution$pm25, col = "light green")
abline(v = 12, col = "red", lwd = 4) # Vertical line at 12
abline(v = median(pollution$pm25), col = "blue", lwd = 4) # Blue line at the median
```

A Barplot is another example, that can be use for categorical data
```{r}
barplot(table(pollution$region), col = "light blue", main = "Number of Counties in Each Region")
```

To get this work you can either create a table with the variable, or build the table into the barplot function